Synthesizing Symmetric Lenses

### 2019-02-19 papers/SymmetricLenses/main.tex
bx proposal changes and make an algorithm more horizontal
add macros and main

\newif\ifdraft\drafttrue  % set true to show comments
% \newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\ifappendices\appendicesfalse

\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
\documentclass[acmsmall,review,screen,anonymous]{acmart}
\settopmatter{printacmref=false,printccs=false}

@@

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

@@

\begin{figure}
  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
  \vspace{-2ex}
  \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
    The user provides regular expressions \BRegex and \BRegexAlt and a set of
    examples \Examples{} as input. \Expand first converts \BRegex and \BRegexAlt
    to stochastic regular expressions \Regex and \RegexAlt with default
    probabilities. It then finds pairs of stochastic regular expressions
    equivalent to \Regex and \RegexAlt and iteratively proposes them to
    \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
    When the algorithm finds a likely lens, it returns it.\bcp{Later: We
      could save significant space if we put the inputs and outputs on the
      sides instead of top and bottom}}
  \label{fig:high-level-algorithm}
\end{figure}

@@

%\begin{enumerate}
%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
%\end{enumerate}
\vskip 1.5ex
\begin{center}
\begin{minipage}[c]{.3\textwidth}
\begin{gather}
 \mbox{$\CreateR \OfType X \rightarrow Y$} \nonumber \\
 \mbox{$\CreateL \OfType Y \rightarrow X$} \nonumber \\
 \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$} \nonumber \\
 \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$} \nonumber
\end{gather} 
\end{minipage} 
\hspace{.4in}
\begin{minipage}[c]{.5\textwidth}
\begin{align}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{align}
\end{minipage} 
\end{center}
\vskip 1ex
\noindent
The two \KW{create} functions are used to propagate shared information and fill
in default values when introducing new data (\EG, on \KW{create} the ``unk''
salary entry is added alongside the name to the management file when HR inserts
a new employee). The two \KW{put} functions propagate edits from one format to
the other by combining a new value from one with an old value from the other.
The notation $\Lens.\PutR$ extracts the \PutR function from simple symmetric
lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
common information between two file types in synch. For example, if a new file
of the left-hand format is created, the \CreateR function will a new file in the
right-hand format. If the right-hand format is edited, the \PutL function can
update the left-hand format with the changed information.

@@

\begin{figure}
  \begin{tabular}{@{\hspace*{5em}}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)\Probability_2}{1 - \Probability_1\Probability_2}})}{\Probability_1\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
  \caption{Stochastic Regular Expression Star-Semiring Equivalence}
  \label{fig:ssr-sre}
\end{figure}

@@

\emph{Stochastic DNF regular expressions} are critical to the success of our
algorithm. Stochastic DNF regular expressions (\SDNFREabbrev{}s) are a
pseudo-normal form for stochastic regular expressions. Of all the star-semiring
equations required to relate two \SDNFREabbrevs, only those relating to Kleene
star (\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs; other axioms, like associativity, commutativity or
distributivity, are unneeded. This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed by the syntax of two
\SDNFREabbrevs, to synthesize any lenses in the same ``class'' (\IE, any lenses
that have the same type when the Kleene star axioms are unused). This property,
and the resulting efficiency of $\GreedySynth$ is key to the effectiveness of
our core algorithms.

>>>>

\newif\ifdraft\drafttrue  % set true to show comments
% \newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anontrue    % set true to suppress names, etc.
\newif\ifappendices\appendicesfalse

\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
%\documentclass[acmsmall,review,screen,anonymous]{acmart}
\documentclass[acmsmall,review,screen,anonymous]{acmart}
\settopmatter{printacmref=false,printccs=false}

@@

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
aboveskip=.3em,
belowskip=.3em,
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

@@

\begin{figure}
  \includegraphics[width=.63\textwidth]{high-level-algorithm.pdf}
  \vspace{-2ex}
  \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
    The user provides regular expressions \BRegex and \BRegexAlt and a set of
    examples \Examples{} as input. \Expand first converts \BRegex and \BRegexAlt
    to stochastic regular expressions \Regex and \RegexAlt with default
    probabilities. It then finds pairs of stochastic regular expressions
    equivalent to \Regex and \RegexAlt and iteratively proposes them to
    \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
    When the algorithm finds a likely lens, it returns it.}
  \label{fig:high-level-algorithm}
\end{figure}

@@

%\begin{enumerate}
%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
%\end{enumerate}
\begin{center}
\begin{minipage}[c]{.3\textwidth}
\begin{gather}
 \mbox{$\CreateR \OfType X \rightarrow Y$} \nonumber \\
 \mbox{$\CreateL \OfType Y \rightarrow X$} \nonumber \\
 \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$} \nonumber \\
 \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$} \nonumber
\end{gather} 
\end{minipage} 
\hspace{.4in}
\begin{minipage}[c]{.5\textwidth}
\begin{align}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x\\
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y\\
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x\\
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{align}
\end{minipage} 
\end{center}
\noindent
\vskip -.5em
The two \KW{create} functions are used to propagate shared information and fill
in default values when introducing new data (\EG, on \KW{create} the ``unk''
salary entry is added alongside the name to the management file when HR inserts
a new employee). The two \KW{put} functions propagate edits from one format to
the other by combining a new value from one with an old value from the other.
The notation $\Lens.\PutR$ extracts the \PutR function from simple symmetric
lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
common information between two file types in synch. For example, if a new file
of the left-hand format is created, the \CreateR function will a new file in the
right-hand format. If the right-hand format is edited, the \PutL function can
update the left-hand format with the changed information.

@@

\begin{figure}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{\hspace*{3em}}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r}
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} &
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} &
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} &
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} &
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} &
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{\hspace{1em}}r@{}}
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)\Probability_2}{1 - \Probability_1\Probability_2}})}{\Probability_1\Probability_2} & \OrAssociativityRule{}  \\
  \end{tabular}
  \caption{Stochastic Regular Expression Star-Semiring Equivalence}
  \label{fig:ssr-sre}
\end{figure}

@@

\emph{Stochastic DNF regular expressions} are critical to the success of our
algorithm. Stochastic DNF regular expressions (\SDNFREabbrev{}s) are a
pseudo-normal form for stochastic regular expressions. Of all the star-semiring
equations required to relate two \SDNFREabbrevs, only those relating to Kleene
star (\EG, \UnrollstarLeftRule{} and \UnrollstarLeftRule{}) are needed to relate
two \SDNFREabbrevs; other axioms, like associativity, commutativity or
distributivity, are unneeded. This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed by the syntax of two
\SDNFREabbrevs, to synthesize any lenses in the same ``class'' (\IE, any lenses
that have the same type when the Kleene star axioms are unused). This property,
and the resulting efficiency of $\GreedySynth$ is key to the effectiveness of
our core algorithms.
